Catchy Title

### 2011-10-11 proposal/paper.tex

SPIRIT \cite{papadimitriou2005streaming} is an algorithm for discovering patterns in multiple streams of timeseries data. This algorithm effectively tries to find correlations between the different data streams by performing principal component analysis on the the data covariance matrix. These principal components are modified and updated as each data sample is received at each time step while forgetting the samples. An interesting aspect of this algorithm is, anomaly detection which is done by detecting the changes in the number of principal components required to capture most of the energy in the signal.
In \cite{roweis1999unifying} the authors describe a unifying model for Linear Gaussian Models which includes PCA, Kalman Filters, Gaussian Mixture Models, ICA etc. They describe a general linear state transition model along with a linear observation model corrupted by independent Gaussian noise. They describe general algorithms to do inference and also learn the parameters of the model.

@@

Another console logs mining paper \cite{xu2009} extracts specific features from unstructured logs of Hadoop system \cite{Hadoop} (A distributed system for data intensive computing) and use them to detect anomalies in the system. The tool is mainly used to detect correctness problems. The author combines logs and source to extract three different signals: message types, state variables, and message identifiers. Based on these signals, they create two matrices: one matrix has state variables as the column, and time as the row, the other matrix uses message types as the column, and message identifiers as the row. Finally, PCA is used for anomaly detection. They perform PCA on the two matrices to get a reduced set of dimensions (eigenvectors) ranked by how much variance in the data each reduced dimension explains. Anomalous rows are identified by flagging those rows whose data is largely unexplained by the first K eigenvectors, which describes 95\% of the overall observed variance. 
After finding the anomalous rows, anomalies need to be manually interpreted to dig out the root causes. In \cite{xu2010}, they further applied their algorithms into some of Google's production logs. Since the data is large which makes it unrealistic to perform PCA, they used frequent mining techniques to segment the long-lasting sequence into sessions.

@@

\subsection{Scalability}
\label{sec:method:scale}
In order to handle multiple streams from large scale data centers, it is necessary to parallelize our analysis algorithm. We plan to use Hadoop as the computing framework to implement the parallel version of our algorithm.

@@

\bibliography{christosref,other}
\bibliographystyle{plain}
\newpage
\pagenumbering{roman}

>>>>

SPIRIT \cite{Papadimitriou2005} is an algorithm for discovering patterns in multiple streams of time-series data. This algorithm effectively tries to find correlations between the different data streams by performing principal component analysis on the the data covariance matrix. These principal components are modified and updated as each data sample is received at each time step while forgetting the samples. An interesting aspect of this algorithm is, anomaly detection which is done by detecting the changes in the number of principal components required to capture most of the energy in the signal.
In \cite{Roweis1999} the authors describe a unifying model for Linear Gaussian Models which includes PCA, Kalman Filters, Gaussian Mixture Models, ICA etc. They describe a general linear state transition model along with a linear observation model corrupted by independent Gaussian noise. They describe general algorithms to do inference and also learn the parameters of the model.

SPIRIT \cite{Papadimitriou2005} is an algorithm for discovering patterns in multiple streams of timeseries data. This algorithm effectively tries to find correlations between the different data streams by performing principal component analysis on the the data covariance matrix. These principal components are modified and updated as each data sample is received at each time step while forgetting the samples. An interesting aspect of this algorithm is, anomaly detection which is done by detecting the changes in the number of principal components required to capture most of the energy in the signal.
In \cite{Roweis1999} the authors describe a unifying model for Linear Gaussian Models which includes PCA, Kalman Filters, Gaussian Mixture Models, ICA etc. They describe a general linear state transition model along with a linear observation model corrupted by independent Gaussian noise. They describe general algorithms to do inference and also learn the parameters of the model.

@@

Another console logs mining paper \cite{xu2009} extracts specific features from unstructured logs of Hadoop system \cite{ApacheSoftwareFoundation2011}  (A distributed system for data intensive computing) and use them to detect anomalies in the system. The tool is mainly used to detect correctness problems. The author combines logs and source to extract three different signals: message types, state variables, and message identifiers. Based on these signals, they create two matrices: one matrix has state variables as the column, and time as the row, the other matrix uses message types as the column, and message identifiers as the row. Finally, PCA is used for anomaly detection. They perform PCA on the two matrices to get a reduced set of dimensions (eigenvectors) ranked by how much variance in the data each reduced dimension explains. Anomalous rows are identified by flagging those rows whose data is largely unexplained by the first K eigenvectors, which describes 95\% of the overall observed variance. 
After finding the anomalous rows, anomalies need to be manually interpreted to dig out the root causes. In \cite{Xu}, they further applied their algorithms into some of Google's production logs. Since the data is large which makes it unrealistic to perform PCA, they used frequent mining techniques to segment the long-lasting sequence into sessions.

@@

\subsection{Scalability}
\label{sec:method:scale}
As a second phase of the project, we plan to choose the model that works best for our data on the basis of how accurately it can provide short-term predictions. As observed above, the primary challenge for many of these algorithms is efficiency and scale. In order to handle multiple streams from large scale data centers, it is necessary to parallelize our analysis algorithm. Though the precise steps we will take to improve algorithm scalability at this point are not yet completely defined, we plan to use Hadoop as the computing framework to implement the parallel version of our algorithm.

@@

\section{Plan of Work}
\begin{tabular}{lll}
\bf Member & \bf Project Portion & \bf Papers Read \\
Yoshihisa Abe & & \cite{Zhu2002}, \cite{Zhu2003}, \cite{Sakurai2011} \\
Ilari Shafer & IPCA application/extension, scalability & \cite{Li2009}, \cite{Gmach2007}, \cite{Lavrenko2000} \\
Kai Ren & Data Collection + Preparation, Anomaly Detection Experiment &  \cite{xu2009}, \cite{Xu}, \cite{nguyen2011} \\
Vishnu Boddeti & Kalman Filter Models & \cite{Papadimitriou2005}, \cite{Roweis1999}
\end{tabular}

\section{Timeline}
\label{sec:timeline}
A rough timeline of the project follows.
\begin{description}
\item[November 10] Finish testing existing models (and preparing data)
\item[November 24] Finish working on distributing algorithm (e.g. Hadoop)
\item[December 4] Finish Final Report
\item[December 9] Address anomaly insertion, clean up loose ends, prepare demo
\end{description}

\bibliography{proposal}
\bibliographystyle{plain}
\newpage
\pagenumbering{roman}
